---
title: "Number of COVID-19 Cases Around the World"
subtitle: "Investigating the number of COVID-19 Cases, Tests, and Deaths by Country, Region, Age, and Sex"
author: 
  - Bruce Zhang
thanks: "Code and data are available at: [https://github.com/RohanAlexander/starter_folder](https://github.com/RohanAlexander/starter_folder)."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(ggplot2)
library(dplyr)
library(here)

# Load the dataset
analysis_data <- read.csv(here("data/02-analysis_data/analysis_data.csv"))

```


# Introduction

Overview paragraph

Estimand paragraph

Results paragraph

Why it matters paragraph

Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....






# Data {#sec-data}

## Overview

We use the statistical programming language R [@citeR] to analyze the data and to create graphs and models. Th packages that were used include tidyverse [@tidyverse], tidyr [@tidyr], dplyr [@dplyr], and rstanarm [@rstanarm]. Our data [@covid_data] was obtained from COVerAGE-DB. The data anlysis was conducted based on the guidance of @tellingstories. 

Overview text

## Measurement
	
Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.

## Outcome variables

Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.



Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

```

Talk way more about it. 

## Predictor variables

Add graphs, tables and text.

Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.








# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.  

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.


### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.


# Results

Our results are summarized in @tbl-modelresults.


```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: cases-by-country
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

# Load necessary library
library(ggplot2)

# Create a bar plot of total cases per country
ggplot(analysis_data, aes(x = Country, y = Cases, fill = Country)) +
  geom_bar(stat = "identity", color = "lightblue", alpha = 0.
)

```

```{r}
#| echo: false
#| eval: true
#| label: cases-over-time
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

# Load necessary libraries
library(ggplot2)
library(dplyr)
library(lubridate) # For date handling

# Filter data to include all regions
filtered_data <- analysis_data %>%
  filter(Region == "All") %>%
  mutate(
    # Ensure the Date column is in proper Date format
    Date = as.Date(Date, format = "%Y-%m-%d"),
    # Extract the year and month to group by
    Month = floor_date(Date, unit = "month")
  )


# Create the scatterplot with aggregated monthly cases
ggplot(filtered_data, aes(x = Date, y = Cases, color = Country, group = Country)) +
  geom_point(alpha = 0.7) +   # Add scatter points
  geom_line(size = 1) +       # Add lines connecting points for each country
  labs(
    title = "Monthly COVID Cases in Canada (Urban Region)",
    x = "Month",
    y = "Total Number of Cases",
    color = "Country"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.minor = element_blank()
  )

```

```{r}
#| echo: false
#| eval: true
#| label: cases-by-age
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

# Load necessary libraries
library(ggplot2)
library(dplyr)

# Filter data for the selected 5 countries
selected_countries <- c("Afghanistan", "Brazil", "Canada", "Denmark", "Egypt")
filtered_data <- analysis_data %>%
  filter(Country %in% selected_countries) %>%
  mutate(
    # Create age intervals
    Age_Interval = cut(Age, breaks = seq(0, 100, by = 10), right = FALSE, include.lowest = TRUE)
  )

# Aggregate the number of cases by country and age interval
age_case_summary <- filtered_data %>%
  group_by(Country, Age_Interval) %>%
  summarise(Total_Cases = sum(Cases, na.rm = TRUE)) %>%
  ungroup()

# Create the faceted histogram
ggplot(age_case_summary, aes(x = Age_Interval, y = Total_Cases, fill = Country)) +
  geom_bar(stat = "identity", alpha = 0.8, color = "black") +
  facet_wrap(~Country, scales = "free_y") + # Create a grid of histograms for each country
  labs(
    title = "COVID Cases by Age Interval in Selected Countries",
    x = "Age Interval",
    y = "Number of Cases",
    fill = "Country"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major = element_line(color = "grey80"),
    legend.position = "none"
  )


```

```{r}
#| echo: false
#| eval: true
#| label: deaths-by-age
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

# Load necessary libraries
library(ggplot2)
library(dplyr)

# Filter data for the selected 5 countries
selected_countries <- c("Afghanistan", "Brazil", "Canada", "Denmark", "Egypt")
filtered_data <- analysis_data %>%
  filter(Country %in% selected_countries) %>%
  mutate(
    # Create age intervals
    Age_Interval = cut(Age, breaks = seq(0, 100, by = 10), right = FALSE, include.lowest = TRUE)
  )

# Aggregate the number of deaths by country and age interval
age_death_summary <- filtered_data %>%
  group_by(Country, Age_Interval) %>%
  summarise(Total_Deaths = sum(Deaths, na.rm = TRUE)) %>% # Use Deaths column
  ungroup()

# Create the faceted histogram for deaths
ggplot(age_death_summary, aes(x = Age_Interval, y = Total_Deaths, fill = Country)) +
  geom_bar(stat = "identity", alpha = 0.8, color = "black") +
  facet_wrap(~Country, scales = "free_y") + # Create a grid of histograms for each country
  labs(
    title = "COVID Deaths by Age Interval in Selected Countries",
    x = "Age Interval",
    y = "Number of Deaths",
    fill = "Country"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major = element_line(color = "grey80"),
    legend.position = "none"
  )

```

```{r}
#| echo: false
#| eval: true
#| label: cases-by-sex
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

# Load necessary libraries
library(ggplot2)
library(dplyr)

# Filter data for the selected 5 countries and group by gender and country
selected_countries <- c("Afghanistan", "Brazil", "Canada", "Denmark", "Egypt")
gender_case_summary <- analysis_data %>%
  filter(Country %in% selected_countries, Sex %in% c("f", "m")) %>% # Filter for selected countries and genders
  group_by(Country, Sex) %>%
  summarise(Total_Cases = sum(Cases, na.rm = TRUE)) %>% # Summarize total cases by gender and country
  ungroup()

# Create the faceted bar graph
ggplot(gender_case_summary, aes(x = Sex, y = Total_Cases, fill = Sex)) +
  geom_bar(stat = "identity", alpha = 0.8, color = "black") + # Bar plot with totals
  facet_wrap(~Country, scales = "free_y") + # Facet by country
  scale_fill_manual(values = c("f" = "pink", "m" = "blue")) + # Optional: Gender-specific colors
  labs(
    title = "COVID Cases by Gender in Selected Countries",
    x = "Gender",
    y = "Number of Cases",
    fill = "Gender"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(size = 12),
    axis.title = element_text(size = 14),
    strip.text = element_text(size = 14),
    panel.spacing = unit(1, "lines") # Increase spacing between facets
  )


```


```{r}
#| echo: false
#| eval: true
#| label: deaths-by-sex
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

# Load necessary libraries
library(ggplot2)
library(dplyr)

# Filter data for the selected 5 countries and group by gender and country
selected_countries <- c("Afghanistan", "Brazil", "Canada", "Denmark", "Egypt")
gender_death_summary <- analysis_data %>%
  filter(Country %in% selected_countries, Sex %in% c("f", "m")) %>% # Filter for selected countries and genders
  group_by(Country, Sex) %>%
  summarise(Total_Deaths = sum(Deaths, na.rm = TRUE)) %>% # Summarize total deaths by gender and country
  ungroup()

# Create the faceted bar graph
ggplot(gender_death_summary, aes(x = Sex, y = Total_Deaths, fill = Sex)) +
  geom_bar(stat = "identity", alpha = 0.8, color = "black") + # Bar plot with totals
  facet_wrap(~Country, scales = "free_y") + # Facet by country
  scale_fill_manual(values = c("f" = "pink", "m" = "blue")) + # Optional: Gender-specific colors
  labs(
    title = "COVID Deaths by Gender in Selected Countries",
    x = "Gender",
    y = "Number of Deaths",
    fill = "Gender"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(size = 12),
    axis.title = element_text(size = 14),
    strip.text = element_text(size = 14),
    panel.spacing = unit(1, "lines") # Increase spacing between facets
  )

```

```{r}
#| echo: false
#| eval: true
#| label: cases-by-region
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

# Load necessary libraries
library(ggplot2)
library(dplyr)

# Filter data for the selected 5 countries and group by region and country
selected_countries <- c("Afghanistan", "Brazil", "Canada", "Denmark", "Egypt")
region_case_summary <- analysis_data %>%
  filter(Country %in% selected_countries) %>% # Filter for selected countries
  group_by(Country, Region) %>%
  summarise(Total_Cases = sum(Cases, na.rm = TRUE)) %>% # Summarize total cases by region and country
  ungroup()

# Create the faceted bar graph
ggplot(region_case_summary, aes(x = Region, y = Total_Cases, fill = Region)) +
  geom_bar(stat = "identity", alpha = 0.8, color = "black") + # Bar plot with totals
  facet_wrap(~Country, scales = "free_y") + # Facet by country
  labs(
    title = "COVID Cases by Region in Selected Countries",
    x = "Region",
    y = "Number of Cases",
    fill = "Region"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 12), # Rotate x-axis labels for readability
    axis.title = element_text(size = 14),
    strip.text = element_text(size = 14),
    panel.spacing = unit(1, "lines") # Increase spacing between facets
  )

```

```{r}
#| echo: false
#| eval: true
#| label: deaths-by-region
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

# Load necessary libraries
library(ggplot2)
library(dplyr)

# Filter data for the selected 5 countries and group by region and country
selected_countries <- c("Afghanistan", "Brazil", "Canada", "Denmark", "Egypt")
region_death_summary <- analysis_data %>%
  filter(Country %in% selected_countries) %>% # Filter for selected countries
  group_by(Country, Region) %>%
  summarise(Total_Deaths = sum(Deaths, na.rm = TRUE)) %>% # Summarize total deaths by region and country
  ungroup()

# Create the faceted bar graph
ggplot(region_death_summary, aes(x = Region, y = Total_Deaths, fill = Region)) +
  geom_bar(stat = "identity", alpha = 0.8, color = "black") + # Bar plot with totals
  facet_wrap(~Country, scales = "free_y") + # Facet by country
  labs(
    title = "COVID Deaths by Region in Selected Countries",
    x = "Region",
    y = "Number of Deaths",
    fill = "Region"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 12), # Rotate x-axis labels for readability
    axis.title = element_text(size = 14),
    strip.text = element_text(size = 14),
    panel.spacing = unit(1, "lines") # Increase spacing between facets
  )

```




# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this. 

## Second discussion point

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}


# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check



```{r}

```

## Diagnostics


```{r}

```



\newpage


# References


